---
title: "A Matter of Taste"
format: html
---

```{r}
library(infer)
library(tidyverse)
```


## Question 1
> List any changes that you made to your experimental protocol from when you formulated it on Wednesday and when you executed it Friday.

We added in random selection of flavors by flipping a coin, assigning heads to lime and tails to lemon, and assigning the treatment based on the results of the coin flip. Before the experiment began, we let the experimenter taste the flavor with knowledge (i.e. not blind) so that they would know what flavor is which. The experimenter rinsed his mouth with water between trials. Added in one straw for each flavored cup as to not cross contaminate. 

## Question 2
> Create a data frame based of the data you collected and print it out into your lab report. You can print all rows your data frame using `
`slice(my_df, 1:100)`. Consult the notes “A Tool for Computing with Data” for a refresher of how to make a data frame.

explanatory column there is a treatment and control
Putting in the data, treatment rows go together, control rows go together 
make use of rep("trt", times = 5)
rep("ctrl", times = 5)
put all of that into one c()
response column yes or no

```{r}
trial <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12)
treatment <- c(TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE)
flavor <- c("Lime", "Lemon", "Lime", "Lime", "Lime", "Lemon", "Lime", "Lemon", "Lemon", "Lime", "Lime", "Lemon")
guess <- c("Lemon", "Lemon", "Lemon", "Lemon", "Lime", "Lime", "Lime", "Lemon", "Lime", "Lime", "Lemon", "Lemon")
data <- data.frame(trial, treatment, flavor, guess)
data <- mutate(data, correct = (flavor == guess))
data
```

## Question ?
> Does the data frame differ at all from the one that you sketched into your experimental protocol? If so, how?

## Question 3
> Create a visualization of the data you collected (not the null distribution) similar to the one you sketched in the handout. Does it look clearly in support of your claim or contrary to your claim or somewhere in between?

```{r}
data %>%
  ggplot(aes(x = treatment, fill = correct)) +
  geom_bar(position = "fill")
```

## Question 4
> Conduct a hypothesis test to determine whether your data is consistent with the null hypothesis. Be sure to provide.  
> a. The null and alternative hypotheses.  
> b. The value of the observed test statistic.  
> c. A visualization of the null distribution and observed test statistic with the p-value shaded in.  
> d. The p-value and your conclusion (use $\alpha$ = 0.05) regarding null hypothesis and original claim.

*H~0~* : There is no difference  
*H~a~* : There is a difference

```{r}
obs_stat <- data %>%
  specify(response = correct, explanatory = treatment, success = "TRUE") %>%
  calculate(stat = "diff in props")

obs_stat
```

```{r}
null <- data %>%
  specify(response = correct, explanatory = treatment, success = "TRUE") %>%
  hypothesize(null = "independence") %>%
  generate(reps = 500, type = "permute") %>%
  calculate(stat = "diff in props") 

null %>%
  visualize() +
  shade_p_value(obs_stat, direction = "both")
```

```{r}
null %>%
  get_p_value(obs_stat, direction = "both")
```

The p-value is greater than the significance level. Therefore, we do not have sufficient evidence to reject the null hypothesis. 

## Question 5
> A thought experiment: if you did not find a significant effect, speculate as to what you could change about your protocol to increase the chance that you find an effect. If you did find a significant effect, speculate as to what you would change about your protocol if you wanted to decrease the chance that you’d find an effect if you were to repeat the experiment.


```{r}

```